---
output:
  pdf_document: default
  html_document: default
---

------------------------------------------------------------------------

title: "Final Project Draft" author: "Florencia Irene" date: "2022-11-17" output: pdf_document

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Cleaning

1.  Missing values

    ```{r}
    led <- read.csv(here::here("data/Life Expectancy Data.csv"))
    ```

    ```{r}
    library(tidyverse)
    library(dplyr)
    library(ggplot2)

    #Pull several variables that are related 
    life <- led %>%
      select(Country, Year, Alcohol, Life.expectancy, Status, Income.composition.of.resources, Schooling, BMI)

    #erase missing values
  life <- na.omit(life)
    


    ```

    | Variable                        | Total Missing Values |
    |---------------------------------|----------------------|
    | Life expectancy                 | 10                   |
    | Alcohol                         | 194                  |
    | Status                          | \-                   |
    | Income composition of resources | 167                  |
    | Schooling                       | 163                  |
    | BMI                             | 34                   |

    Everything is below 10%, so it is acceptable.

2.  Rename variable & Factor levels

    ```{r}
    led <- led %>%
      rename(Education = Schooling, 
             ICOR = Income.composition.of.resources)%>%
      mutate(Status = factor(Status))
    ```

## Exploratory Data Analysis

1.  Summary Statistic

    ```{r}
    summary(led)
    ```

2.  Conditions

    1.  Scatter Plot

    ```{r}
    life %>%
    ggplot(aes(x = Alcohol, y =Life.expectancy, color = Alcohol)) +
      geom_point(alpha =0.3 , shape = 16, size = 2) +
      theme_minimal() +
      scale_color_gradient(low = "#0091ff", high = "#f0650e")+ 
      geom_smooth(method = "lm",se= FALSE, show.legend =  FALSE) +
      labs(title = "Relationship of Life Expectancy and Alcohol", 
           x = "Alcohol Consumption (in litres of pure alcohol)",
           y = "Life expectancy(in age)")

    ```


2.  Normality

```{r}
#check the normality
y <- life$Life.expectancy
x <- life$Alcohol

mod <- lm(y ~ x, data = life)

qqnorm(resid(mod))
qqline(resid(mod), col = "red")



```

3.  Equal variance of residuals

```{r}
#checking constant variance
plot(resid(mod)~fitted(mod), main = "Residuals vs fitted")
abline(h = 0, col = "red")
```

3.  Result

    ```{r}
    summary(mod)

    #p value
    summary(mod)$coefficients[,4] 

    #calculating p value
    pval <- 1- 9.418124e-89/2
    pval

    #confidence interval
    confint(mod)

    #correlation 
    cor(x, y)
    ```

**Based on ICOR**

1.  Plot

```{r}
#install gghighlight 
install.packages("gghighlight")
library(gghighlight)

#Plot highlight low icor (<0.5)
life %>%
  ggplot(aes(
    x = Alcohol, 
    y = Life.expectancy, 
    color = ICOR))+
  geom_point( alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Life Expectancy, Alcohol, and ICOR", 
       x = "Alcohol Consumption (in litres of pure alcohol)",
       y = "Life expectancy(in age)")+
  theme_minimal()+scale_color_viridis_b()+
  gghighlight(ICOR < 0.550)

```

```{r}
#Plot of Alcohol, life expectancy based on ICOR
life %>%
  ggplot(aes(
    x = Alcohol, 
    y = Life.expectancy, 
    color = ICOR))+
  geom_point( alpha = 0.2) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Life Expectancy, Alcohol, and ICOR", 
       x = "Alcohol Consumption (in litres of pure alcohol)",
       y = "Life expectancy(in age)")+
  theme_minimal()+scale_color_viridis_b()
```

```{r}
#summary statistic of alcohol consumption and life expectancy based on ICOR
life %>%
  select(Alcohol, Life.expectancy, ICOR) %>% 
  filter (ICOR >= 0.5)%>%
  summarise (mean_alcohol = mean(Alcohol),
             sd_Alcohol = sd(Alcohol),
             mean_life.expectancy = mean(Life.expectancy),
             sd_life.expectancy = sd(Life.expectancy))

life %>%
  select(Alcohol, Life.expectancy, ICOR) %>% 
  filter (ICOR < 0.5)%>%
  summarise (mean_alcohol = mean(Alcohol),
             sd_Alcohol = sd(Alcohol),
             mean_life.expectancy = mean(Life.expectancy),
             sd_life.expectancy = sd(Life.expectancy))

```
